Abstract Binding Trees

### 2015-11-07 paper.tex
[paper] Rough implementation of the interpretation
@freebroccolo This seems like it should work; could you give me a second
pair of eyes, though? Incidentally, I think that the factoring I have
proposed here will make the Agda proof more elegant if we should adopt
it.

This probably should be reorganized a bit so it's not just a block of
definitions(?), but this is what "hole-driven-development" in LaTeX
looks like :)

For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
endofunctor
$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\to\widehat{\HypCat}^\Sorts}$,
which is defined as follows:
\begin{align*}
  \IMode{{\SigFun{\Sigma}(X)}_{\tau}}
    &\triangleq
    \OMode{
      \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}
      \prod_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau'}\in\vec{v}}
      X_{\tau'}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
    }
\end{align*}

@@

Next we have the interpretation of metavariable applications
\[%
  \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
    \MApp{\MV{m}}{\vec{u}}{\vec{M}} : \tau}_P%
  \qquad(\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}})
\]
which proceeds by projecting the metavariable's interpretation from the
environment and intantiating it via substitution:
\[
  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
          \arrow[r, "\IMode{\Tuple{\pi_\MV{m}\,\pi_1, \phi\,\pi_2, \psi}}"]
    \& {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
          \arrow[r, "\IMode{\varsigma_{\vec{\sigma}\parallel\vec{\tau}}}"]
    \& \IMode{P_s}
  \end{tikzcd}
  \tag{metavar}
\]
where
\begin{align*}
  \ADefine{\phi}{
  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
    {S^{\{\Upsilon\}}} \arrow[r, "\Tuple{\pi_u}_{u\in\vec{u}}"]
    \& {S^{\{\vec{\sigma}\}}}
  \end{tikzcd}
  }\\
  \ADefine{\psi}{
   \begin{tikzcd}[ampersand replacement = \&, sep=28ex]
     {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}
     \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}"]
     \& {P^{[\vec{\tau}]}}
   \end{tikzcd}
  }
\end{align*}

The interpretation of operator applications is most complicated. Recall that,
unlike in standard treatments of algebraic theories, our operators are indexed
by symbol collections; therefore, operators must pass through suitable
renamings in order to be used in the interpretation. Let us begin by
constructing for each operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the
natural transformation
$\Of{\SemBrackets{\vartheta}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
\to\Operators_{a}}$, which shall rename the parameters of the operator using
the environment.

@@

>>>>

For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
endofunctor
$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\to\widehat{\HypCat}^\Sorts}$,
which is defined as follows:
\begin{align*}
  \IMode{{\SigFun{\Sigma}(X)}_{\tau}}
    &\triangleq
    \OMode{
      \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}
      \prod_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau_i}\in\vec{v}}
      X_{\tau_i}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}
    }
\end{align*}

@@

Next we have the interpretation of metavariable applications
\[%
  \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash%
    \MApp{\MV{m}}{\vec{u}}{\vec{M}} : \tau}_P%
  \qquad(\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}})
\]
which proceeds by projecting the metavariable's interpretation from the
environment and intantiating it via substitution:
\[
  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}}
          \arrow[r, "\IMode{\Tuple{\pi_\MV{m}\,\pi_1, \phi\,\pi_2, \psi}}"]
    \& {\IMode{P_{\tau}^{\Yoneda{\vec{\sigma}\parallel\vec{\tau}}}\times S^{\{\vec{\sigma}\}} \times P^{[\vec{\tau}]}}}
          \arrow[r, "\IMode{\varsigma_{\vec{\sigma}\parallel\vec{\tau}}}"]
    \& \IMode{P_\tau}
  \end{tikzcd}
  \tag{metavar}
\]
where
\begin{align*}
  \ADefine{\phi}{
  \begin{tikzcd}[ampersand replacement = \&, sep=huge]
    {S^{\{\Upsilon\}}} \arrow[r, "\Tuple{\pi_u}_{u\in\vec{u}}"]
    \& {S^{\{\vec{\sigma}\}}}
  \end{tikzcd}
  }\\
  \ADefine{\psi}{
   \begin{tikzcd}[ampersand replacement = \&, sep=28ex]
     {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}
     \arrow[r, "\Tuple{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M:\tau}_P}_{(M,\tau)\in (\vec{M},\vec{\tau})}"]
     \& {P^{[\vec{\tau}]}}
   \end{tikzcd}
  }
\end{align*}

The interpretation of operator applications is most complicated. Recall that,
unlike in standard treatments of universal algebra, our operators are indexed
by symbol collections; therefore, operators must pass through suitable
renamings in order to be used in the interpretation. Let us begin by
constructing for each operator $\IsOperator{\Upsilon}{\vartheta}{a}$ the
natural transformation
$\Of{\SemBrackets{\vartheta}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
\to\Operators_{a}}$, which shall rename the parameters of the operator using
the environment.

@@

We will proceed using the initial $\SigFun{\Sigma}$-algebra $\alpha$, as
follows.
\[
   \begin{tikzcd}[ampersand replacement = \&, sep=large]
     {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\OMode{\beta}"]
     \& {\IMode{\SigFun{\Sigma}(P)_\tau}} \arrow[r, "\IMode{\alpha}"]
     \& {\IMode{P_\tau}}
   \end{tikzcd}
\]

\begin{align*}
  \ADefine{
    \beta
  }{
   \begin{tikzcd}[ampersand replacement = \&, sep=huge]
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\Tuple{\SemBrackets{\vartheta}_P, \gamma}"]
      \& {
        \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{\tau}}}
        \prod_{\MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}}
        P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
      }
   \end{tikzcd}
  }\\
  \ADefine{
    \gamma
  }{
   \begin{tikzcd}[ampersand replacement = \&, sep=6em]
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\Tuple{\lambda x. \eta_i(-,x))}_i"]
      \& {
        \prod_{\MkValence{\vec{\sigma}_i}{\vec{\tau}_i}{\tau_i}\in\vec{v}}
        P_{\tau_i}^{\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_i}}
      }
   \end{tikzcd}
  }\\
  \ADefine{
    \eta_i
  }{
   \begin{tikzcd}[ampersand replacement = \&, sep=7em]
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
      } \arrow[r, "\Tuple{\pi_1\circ\pi_1,\phi_i,\nu_{\Gamma,\vec{\tau}_i}\circ\psi_i}"]
      \& {\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i}_P}
      \arrow[r, "\SemBrackets{\Omega\triangleright\Upsilon,\vec{\sigma}_i\parallel\Gamma,\vec{\tau}_i\vdash M_i:\tau_i}_P"]
      \& {P_{\tau_i}}
   \end{tikzcd}
  }\\
  \ADefine{
    \phi_i
  }{
   \begin{tikzcd}[ampersand replacement = \&, sep=12em]
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
      } \arrow[r, "\Tuple{\pi_u\circ\pi_1}_{u\in\Dom\Upsilon}\oplus\Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}"]
      \& {S^{\{\Upsilon,\vec{\sigma}_i\}}}
   \end{tikzcd}
  }\\
  \ADefine{
    \psi_i
  }{
   \begin{tikzcd}[ampersand replacement = \&, sep=12em]
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P
        \times\Yoneda{\vec{\sigma}_i\parallel\vec{\tau}_u}
      } \arrow[r, "\Tuple{\pi_u\circ\pi_1}_{u\in\Dom\Upsilon}\oplus\Tuple{\pi_u\circ\pi_1\circ\pi_2}_{u\in\Dom{\vec{\sigma}_i}}"]
      \& {V^{[\Gamma,\vec{\tau}_i]}}
   \end{tikzcd}
  }
\end{align*}
